---
title: "Carbon Cities"
format: 
  html:
    theme: cosmo
    toc: true
    code-fold: true
    self-contained: false 
    page-layout: full
editor: visual
warning: false
date: 04/05/2026
date-format: long
author: 
 - name: He Yutian
abstract: > 
  A study of 53 C40 cities, as part of "Speculative Carbon Futures Amidst Climate Data Inequities (working title)"
---

## Grid pixel heatmap demo (dash plotly)

-   Data exploration to identify trends; shows spatial-temporal spread of data, data gaps, and allows for more detailed sorting

-   Inverse relationship between CO2 and SIF - hypothesis about SIF's impact on CO2? Nature-based solutions

```{=html}
<iframe src="http://localhost:8052" height="700" width="100%"></iframe>

<!--<iframe src="http://localhost:8052" height="700" width="100%" style="border:none; pointer-events: none;"></iframe>-->-->
```

## Visual Prototype Overview (Figma)

-   Building a tool to go from location-based selection of points of interest to comparing variables relevant to CO2

```{=html}
<iframe style="border: 1px solid rgba(0, 0, 0, 0.1);" width="100%" height="700" src="https://embed.figma.com/proto/nOUZp57ZbuHiR0M840E261/OCO3?node-id=553-4205&p=f&scaling=contain&content-scaling=fixed&page-id=0%3A1&starting-point-node-id=553%3A4205&show-proto-sidebar=1&embed-host=share" allowfullscreen></iframe>
```

## Visual prototype demo (js)

-   Demo of some of the Figma prototype functionalities using real data; WIP.

```{=html}
<iframe src="globe/index.html" width="100%" height="900px" style="border:none;"></iframe>
```

## Regression analysis (53 cities)

Carbon impact score (CO2 conc - CO2 emissions)

-   A 'carbon impact score' is introduced by taking the normalized interquartile range of CO2 concentration levels (OCO-3 netcdf data) subtracted by CO2 emissions levels (satellite-based estimates from Ahn et al, 2025).

-   Higher score = more "harmed" (high conc low emissions), lower score = more "harming" done (low conc high emissions)

-   This is then compared against solar induced fluorescence (SIF), population, gdp, number of samples, and coastal proximity via running a multivariate linear regression.

```{r}
# reference: https://towardsdatascience.com/7-steps-to-run-a-linear-regression-analysis-using-r-d5897a66b835/

# install.packages("tidyverse")
library(tidyverse)


df <- read.csv("c40cities-co2-em-conc-wkt-coast.csv") # outcome of https://github.com/heytian/d2d-oco3-tools/blob/main/20260405_04_wkt-coastal.py
df <- df |> drop_na(population)

# normalize independent variables using interquartile range 
iqr_norm <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  (x - q1) / (q3 - q1)
}


norm_df <- df |> 
  mutate(
    co2_conc_iqr = iqr_norm(xco2_ppm),
    co2_em_iqr = iqr_norm(annual_co2_em_mtco2),
    sif_iqr = iqr_norm(sif_757nm),
    pop_iqr = iqr_norm(population),
    xco2numsams_iqr = iqr_norm(xco2_numsams),
    sifnumsams_iqr = iqr_norm(sif_numsams),
    gdp_iqr = iqr_norm(gdp_bil_usd),
    coast_iqr = iqr_norm(coastal_km),
    impact_iqr = co2_conc_iqr - co2_em_iqr
  )
    

lm_iqr_impact<- lm(impact_iqr ~ sif_iqr + pop_iqr + xco2numsams_iqr + sifnumsams_iqr + gdp_iqr + coast_iqr, 
                  data = norm_df)
summary(lm_iqr_impact)
```

## Plots

This is the distribution of cities across the quadrant framework of high-low CO2 concentrations and emissions.

```{r}
# bubble size = population, color = quadrant, shape = coastal/inland
ggplot(norm_df, aes(x = co2_em_iqr, y = co2_conc_iqr)) +
  
  # quadrant dividing lines at IQR thresholds (0 = Q1, 1 = Q3)
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray60") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray60") +
  
  # cities as bubbles
  geom_point(aes(
    color = quadrant,
    size  = population / 1e6, # size by population millions
    shape = is_coastal
  ), alpha = 0.75) +
  
  # label interesting cities only
  ggrepel::geom_text_repel(
    data = norm_df |> filter(
      quadrant %in% c("LOW_emissions_HIGH_conc", "HIGH_emissions_LOW_conc")
    ),
    aes(label = city),
    size = 2.8, max.overlaps = 15
  ) +
  
  scale_color_manual(values = c(
    "HIGH_both"              = "#d73027",
    "LOW_both"               = "#4575b4",
    "HIGH_emissions_LOW_conc" = "#fc8d59",
    "LOW_emissions_HIGH_conc" = "#91bfdb")) +
  scale_shape_manual(values = c("True" = 17, "False" = 16),
                     labels = c("True" = "Coastal", "False" = "Inland")) +
  scale_size_continuous(range = c(2, 10), name = "Population (M)") +
  
  labs(
    title    = "City CO2 Concentration vs Emissions",
    subtitle = "Quadrants based on IQR thresholds | Size = population | Shape = coastal/inland",
    x        = "CO2 Emissions (IQR normalized)",
    y        = "CO2 Concentration (IQR normalized)",
    color    = "Quadrant",
    shape    = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

Impact score by quadrant, color differentiated by coastal proximity. Coastal cities cluster differently across the CO2 emissions - CO2 concentration quadrants

```{r}
ggplot(norm_df, aes(x = reorder(quadrant, impact_iqr, median), 
                     y = impact_iqr)) +
  geom_boxplot(outlier.shape = NA, fill = "gray90") +
  geom_jitter(aes(color = coastal_km, size = population / 1e6),
              width = 0.2, alpha = 0.8) +
  scale_color_viridis_c(
    name   = "Distance to\ncoast (km)",
    option = "viridis"
  ) +
  scale_size_continuous(range = c(2, 6), name = "Population (M)") +
  labs(
    title    = "Impact Score Distribution by Quadrant",
    subtitle = "Color = coastal distance | Size = population",
    x        = NULL,
    y        = "Impact Score (IQR normalized)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

Regression analysis shows that coastal distance and population are key drivers.

```{r}
library(tidyverse)
library(broom)

# tidy regression output
coef_df <- tidy(lm_iqr_impact, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(term = recode(term,
    "sif_iqr"         = "SIF",
    "pop_iqr"         = "Population",
    "xco2numsams_iqr" = "CO2 Sample Size",
    "sifnumsams_iqr"  = "SIF Sample Size",
    "gdp_iqr"         = "GDP",
    "coast_iqr"       = "Coastal Distance"
  ))

ggplot(coef_df, aes(x = estimate, y = reorder(term, estimate))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(aes(color = p.value < 0.20), size = 3) +
  scale_color_manual(
    values = c("FALSE" = "gray70", "TRUE" = "red"),
    labels = c("p > 0.20", "p < 0.20")
  ) +
  labs(
    title = "Predictors of CO2 Impact Score",
    subtitle = "Impact score = normalized CO2 concentration − normalized emissions",
    x = "Coefficient estimate (IQR normalized)",
    y = NULL,
    color = NULL
  ) +
  theme_minimal()
```

The coastal effect is most visible in cities with low CO2 emissions and high CO2 concentrations. Coastal proximity may partially explain why these cities end up receiving more CO2 despite emitting low amounts.

```{r}
ggplot(norm_df, aes(x = coastal_km, y = impact_iqr)) +
  geom_point(aes(color = quadrant, size = population / 1e6), alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "gray30", linewidth = 0.7) +
  scale_color_manual(values = c(
    "HIGH_both"               = "#d73027",
    "LOW_both"                = "#4575b4",
    "HIGH_emissions_LOW_conc" = "#fc8d59",
    "LOW_emissions_HIGH_conc" = "#91bfdb"
  )) +
  scale_size_continuous(range = c(2, 8), name = "Population (M)") +
  facet_wrap(~quadrant, scales = "free_y") +
  labs(
    title    = "Coastal Distance vs Impact Score by Quadrant",
    x        = "Distance to coast (km)",
    y        = "Impact Score (IQR normalized)",
    color    = "Quadrant"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Regression analysis (346 locations)

WIP (Apr 8)

Compare parquet files with location csv to check for SAM target coverage:

```{r}
library(arrow)
library(dplyr)
library(lubridate)
library(tidyr)
locations <- read.csv("clasp_report_356sams-pop-gdp-coast.csv")

# replace parquet paths with local paths
oco3co2 <- read_parquet("/Users/heyutian/Downloads/oco3_parquet/co2_sam.parquet")
oco3sif <- read_parquet("/Users/heyutian/Downloads/oco3_parquet/sif_sam.parquet")
oco3co2sif <- bind_rows(oco3co2, oco3sif)

# diagnostic: check for overlaps between target names in the 2 files
target_name_csv <- unique(locations$Target.Name)
target_name_parquet <- unique(oco3co2sif$target_name)
missing_in_csv <- setdiff(target_name_parquet, target_name_csv)
missing_in_parquet <- setdiff(target_name_csv, target_name_parquet)
cat("Parquet entries NOT in CSV:\n")
print(missing_in_csv)
cat("\nCSV entries NOT in parquet:\n")
print(missing_in_parquet)
```

Check variables in both files

```{r}
colnames(locations)
colnames(oco3co2sif)
```

Merge/ analyze

```{r}
locations_select <- locations |>
  select(
    target_name = "Target.Name",  
    wkt_center="Site.Center.WKT",
    wkt_shape="Site.Shape.WKT",
    city,
    country,
    population,
    gdp_bil_usd,
    coastal_km,
    is_coastal
  )

oco3_full <- oco3co2sif |>
  select(-any_of(c("city","country","population"))) |>
  left_join(locations_select, by = "target_name")

print(colnames(oco3_full))
print(sum(!is.na(oco3_full$country)))

```

```{r}
oco3_full <- oco3_full |>
  mutate(
    datetime = as.POSIXct(datetime),
    year = year(datetime),
    month = month(datetime),
    day_of_year = yday(datetime),
    season = case_when(
      latitude >= -20 & latitude <= 20 & month >= 5 & month <= 10 ~ "Tropical Wet",
      latitude >= -20 & latitude <= 20 ~ "Tropical Dry",
      
      latitude > 20 & latitude <= 60 & month %in% c(12, 1, 2) ~ "Winter (North)",
      latitude > 20 & latitude <= 60 & month %in% c(3, 4, 5) ~ "Spring (North)",
      latitude > 20 & latitude <= 60 & month %in% c(6, 7, 8) ~ "Summer (North)",
      latitude > 20 & latitude <= 60 & month %in% c(9, 10, 11) ~ "Autumn (North)",
      
      latitude >= -60 & latitude < -20 & month %in% c(12, 1, 2) ~ "Summer (South)",
      latitude >= -60 & latitude < -20 & month %in% c(3, 4, 5) ~ "Autumn (South)",
      latitude >= -60 & latitude < -20 & month %in% c(6, 7, 8) ~ "Winter (South)",
      latitude >= -60 & latitude < -20 & month %in% c(9, 10, 11) ~ "Spring (South)",
      
      TRUE ~ "Other"
    ),
    hour = hour(local_time),
    time_of_day = case_when(
      hour < 6  ~ "Night",
      hour < 12 ~ "Morning",
      hour < 18 ~ "Afternoon",
      TRUE ~ "Evening"
    )
  )

iqr_norm <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  (x - q1) / (q3 - q1)
}

regression_df <- oco3_full |>
  group_by(target_name, country, population, gdp_bil_usd, coastal_km, season, time_of_day) |>
  summarise(
    city = first(city),
    country = first(country),
    population = first(population),
    gdp_bil_usd = first(gdp_bil_usd),
    coastal_km = first(coastal_km),
    xco2_median = median(xco2, na.rm = TRUE),
    xco2_std = sd(xco2, na.rm = TRUE),
    xco2_n = n(),
    sif_median = median(Daily_SIF_757nm, na.rm = TRUE),
    sif_std = sd(Daily_SIF_757nm, na.rm = TRUE),
    sif_n = n(),
    .groups = "drop"
  ) |>
  filter(!is.na(population)) |>
  mutate(
    xco2_iqr = iqr_norm(xco2_median),
    sif_iqr = iqr_norm(sif_median),
    pop_iqr = iqr_norm(population),
    xco2_numsams_iqr = iqr_norm(xco2_n),
    sif_numsams_iqr = iqr_norm(sif_n),
    gdp_iqr = iqr_norm(gdp_bil_usd),
    coast_iqr = iqr_norm(coastal_km),
    season = as.factor(season),
    time_of_day = as.factor(time_of_day)
  ) 
```

```{r}
model_interactions <- lm(xco2_iqr ~ (sif_iqr + pop_iqr + xco2_numsams_iqr + gdp_iqr + coast_iqr) * season, 
                         data = regression_df)

summary(model_interactions)
```

```{r}

```

```{r}
# Separate models for each season
seasons_list <- unique(regression_df$season)

for (s in seasons_list) {
  cat("\n=== Season:", s, "===\n")

  model_s <- lm(xco2_iqr ~ sif_iqr + pop_iqr + xco2_numsams_iqr + gdp_iqr + coast_iqr,
                data = regression_df %>% filter(season == s))

  print(summary(model_s))
}
```

Findings:

|  |  |
|--------------------------|----------------------------------------------|
| **Seasonal variation** | Overall R² = 0.156; interactions explain 15.6% of variance |
| **Summer (North) is strongest** | R² = 0.095, p \< 0.001; SIF, pop, GDP, coast all significant |
| **Winter (South) second strongest** | R² = 0.178, coastal effect significant (p = 0.018) |
| **SIF consistently matters** | Negative across most seasons (higher vegetation = lower CO2 anomalies) |
| **Coastal distance is seasonal** | Critical in Summer/Winter North; less relevant in tropics |
| **Tropics are unpredictable** | R² \< 0.05 for Tropical Dry/Wet; weak relationships |
| **Southern Hemisphere weaker** | Generally lower R² than Northern Hemisphere |

```{r}

summary(lm(xco2_median ~ sif_median + pop_iqr + gdp_iqr + coast_iqr,
   data = regression_df %>% filter(season == "Summer (North)")))
```

Findings:

**SIF: -6.42 (p\<0.001)**\*
Vegetation draws CO2 down strongly. Sites with suppressed SIF — urban, industrial, degraded land — carry higher atmospheric CO2.

**Population: +0.44 (p=0.019)**
Denser populations bear higher XCO2.

**GDP: -0.25 (p=0.038)**
Wealthier sites have lower XCO2 median.

**Coastal: -0.83 (p\<0.001)**
Ocean sink proximity reduces XCO2.

|     |     |
|-----|-----|
|     |     |
|     |     |
|     |     |
|     |     |
|     |     |
|     |     |
|     |     |

```{r}
library(ggplot2)
library(gridExtra)

season_models <- data.frame(
  season = c("Tropical Dry", "Tropical Wet", "Autumn (South)", "Spring (South)", 
             "Summer (South)", "Winter (South)", "Autumn (North)", "Spring (North)", 
             "Summer (North)", "Winter (North)"),
  r_squared = c(0.032, 0.025, 0.036, 0.054, 0.105, 0.178, 0.029, 0.038, 0.095, 0.018),
  sif_sig = c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, TRUE, FALSE),
  coast_sig = c(FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, TRUE, FALSE),
  hemisphere = c("Tropical", "Tropical", "South", "South", "South", "South", "North", "North", "North", "North"),
  p_value = c(0.606, 0.635, 0.713, 0.446, 0.111, 0.037, 0.005, 0.0002, 0.0000, 0.171)
)

p1 <- ggplot(season_models, aes(x = r_squared, y = -log10(p_value), 
                                 color = hemisphere, size = r_squared)) +
  geom_point(alpha = 0.6, pch = 19) +
  geom_text(aes(label = substr(season, 1, 10)), nudge_y = 0.15, size = 3) +
  geom_vline(xintercept = 0.05, linetype = "dashed", alpha = 0.5) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c("North" = "#FF6B6B", "South" = "#4ECDC4", "Tropical" = "#FFE66D")) +
  labs(
    title = "Season Effectiveness: R² vs Statistical Significance",
    x = "R-squared (Model Fit)",
    y = "-log10(p-value)",
    color = "Hemisphere",
    size = "R²"
  ) +
  theme_minimal() +
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5, face = "bold"))

print(p1)

coef_comparison <- data.frame(
  season = c("Summer (North)", "Winter (North)", "Spring (North)", "Autumn (North)",
             "Summer (South)", "Winter (South)", "Spring (South)", "Autumn (South)",
             "Tropical Dry", "Tropical Wet"),
  sif_coef = c(-0.205, -0.282, -0.072, -0.258,
               -0.193, 0.290, -0.072, -0.108,
               -0.015, -0.117),
  coast_coef = c(-0.134, -0.041, 0.015, 0.017,
                 0.259, 0.435, -0.098, 0.208,
                 -0.140, -0.015),
  hemisphere = c("North", "North", "North", "North", 
                 "South", "South", "South", "South",
                 "Tropical", "Tropical")
)

p2 <- ggplot(coef_comparison, aes(x = sif_coef, y = coast_coef, color = hemisphere)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_text(aes(label = substr(season, 1, 8)), nudge_x = 0.02, size = 2.5) +
  geom_hline(yintercept = 0, linetype = "solid", alpha = 0.3) +
  geom_vline(xintercept = 0, linetype = "solid", alpha = 0.3) +
  scale_color_manual(values = c("North" = "#FF6B6B", "South" = "#4ECDC4", "Tropical" = "#FFE66D")) +
  labs(
    title = "Quadrant: SIF Effect vs Coastal Distance Effect",
    x = "SIF Coefficient (← vegetation helps)",
    y = "Coastal Coefficient (↑ coast increases CO2)",
    color = "Hemisphere"
  ) +
  annotate("text", x = -0.25, y = 0.4, label = "SIF dominant\nCoast positive", size = 3, alpha = 0.6) +
  annotate("text", x = -0.25, y = -0.1, label = "Both reduce\nCO2", size = 3, alpha = 0.6) +
  theme_minimal() +
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5, face = "bold"))

print(p2)

p3 <- ggplot(season_models, aes(x = reorder(season, r_squared), y = r_squared, 
                                 fill = ifelse(p_value < 0.05, "Sig", "Not Sig"))) +
  geom_col(alpha = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("Sig" = "#2ECC71", "Not Sig" = "#BDC3C7")) +
  labs(
    title = "Model R² by Season\n(Green = Significant, p < 0.05)",
    x = "Season",
    y = "R-squared",
    fill = "Significance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

print(p3)

predictor_sig <- data.frame(
  season = rep(c("Summer (N)", "Winter (N)", "Spring (N)", "Winter (S)"), 4),
  predictor = c(rep("SIF", 4), rep("Population", 4), rep("Coast", 4), rep("xco2_numsams", 4)),
  significant = c(
    TRUE, FALSE, FALSE, FALSE,  # SIF
    TRUE, FALSE, TRUE, FALSE,   # Population
    TRUE, FALSE, FALSE, TRUE,   # Coast
    FALSE, FALSE, TRUE, FALSE   # xco2_numsams
  )
)

p4 <- ggplot(predictor_sig, aes(x = predictor, y = season, fill = significant)) +
  geom_tile(color = "white", size = 1) +
  scale_fill_manual(values = c("TRUE" = "#3498DB", "FALSE" = "#ECF0F1")) +
  labs(
    title = "Significant Predictors by Season\n(Blue = p < 0.05)",
    x = "Predictor",
    y = "Season",
    fill = "Significant"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

print(p4)
```

```{r}
library(ggplot2)
library(ggrepel)

# Create quadrants based on percentiles
viz_data <- regression_df %>%
  # filter(city != "Urumqi") %>%
  mutate(
    sif_threshold = median(sif_iqr, na.rm = TRUE),
    co2_threshold = median(xco2_iqr, na.rm = TRUE),
    
    quadrant = case_when(
      sif_iqr > sif_threshold & xco2_iqr > co2_threshold ~ "High SIF\nHigh CO2",
      sif_iqr > sif_threshold & xco2_iqr <= co2_threshold ~ "High SIF\nLow CO2",
      sif_iqr <= sif_threshold & xco2_iqr > co2_threshold ~ "Low SIF\nHigh CO2",
      sif_iqr <= sif_threshold & xco2_iqr <= co2_threshold ~ "Low SIF\nLow CO2"
    ),
    
    coastal_status = if_else(coastal_km < 50, "Coastal", "Inland")
  )

# Only keep extreme cities
extreme_data <- viz_data %>%
  group_by(quadrant) %>%
  arrange(desc(abs(sif_iqr) + abs(xco2_iqr))) %>%  # Distance from center
  slice_head(n = 6) %>%
  ungroup()

# Plot
ggplot(viz_data, aes(x = sif_iqr, y = xco2_iqr)) +
  
  # Quadrant dividing lines
  geom_hline(yintercept = median(viz_data$xco2_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.8, alpha = 0.7) +
  geom_vline(xintercept = median(viz_data$sif_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.8, alpha = 0.7) +
  
  # Background all cities (small, gray)
  geom_point(size = 2, color = "lightgray", alpha = 0.3) +
  
  # Highlight extreme cities
  geom_point(
    data = extreme_data,
    aes(color = quadrant, size = population / 1e6, shape = coastal_status),
    alpha = 0.8,
    stroke = 1
  ) +
  
  # Label extreme cities only
  ggrepel::geom_text_repel(
    data = extreme_data,
    aes(label = city, color = quadrant),
    size = 3.5,
    max.overlaps = 20,
    segment.size = 0.4,
    segment.alpha = 0.5,
    box.padding = 0.4,
    fontface = "bold"
  ) +
  
  scale_color_manual(
    values = c(
      "High SIF\nHigh CO2" = "#e41a1c",   # Red - unexpected
      "High SIF\nLow CO2" = "#4daf4a",    # Green - good
      "Low SIF\nHigh CO2" = "#ff7f00",    # Orange - bad
      "Low SIF\nLow CO2" = "#377eb8"      # Blue - neutral
    ),
    name = "Quadrant"
  ) +
  
  scale_shape_manual(
    values = c("Coastal" = 17, "Inland" = 16),
    name = "Type"
  ) +
  
  scale_size_continuous(
    range = c(2, 5),
    name = "Population (M)"
  ) +
  
  facet_wrap(~season, ncol = 2) +
  
  labs(
    title = "Extreme Cities: SIF Impact vs CO2 Concentration by Season",
    subtitle = "Showing 6 extreme cities per quadrant | ▲ = Coastal | ● = Inland | Gray dots = all other cities",
    x = "SIF Effect (← helps reduce CO2)",
    y = "CO2 Concentration"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    strip.text = element_text(face = "bold", size = 9),
    panel.grid.minor = element_blank(),
    aspect.ratio = 0.5
  )

options(repr.plot.width = 16, repr.plot.height = 12)
```

```{r}
library(ggplot2)
library(ggrepel)

# Create quadrants based on percentiles
viz_data <- regression_df %>%
  mutate(
    sif_threshold = median(sif_iqr, na.rm = TRUE),
    co2_threshold = median(xco2_iqr, na.rm = TRUE),
    
    quadrant = case_when(
      sif_iqr > sif_threshold & xco2_iqr > co2_threshold ~ "High SIF\nHigh CO2",
      sif_iqr > sif_threshold & xco2_iqr <= co2_threshold ~ "High SIF\nLow CO2",
      sif_iqr <= sif_threshold & xco2_iqr > co2_threshold ~ "Low SIF\nHigh CO2",
      sif_iqr <= sif_threshold & xco2_iqr <= co2_threshold ~ "Low SIF\nLow CO2"
    ),
    
    coastal_status = if_else(coastal_km < 50, "Coastal", "Inland")
  )

# Only keep extreme cities
extreme_data <- viz_data %>%
  group_by(quadrant) %>%
  arrange(desc(abs(sif_iqr) + abs(xco2_iqr))) %>%
  slice_head(n = 6) %>%
  ungroup()

# ============== PLOT 1: ALL SEASONS COMBINED ==============
p_combined <- ggplot(viz_data, aes(x = sif_iqr, y = xco2_iqr)) +
  
  # Quadrant dividing lines
  geom_hline(yintercept = median(viz_data$xco2_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.8, alpha = 0.7) +
  geom_vline(xintercept = median(viz_data$sif_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.8, alpha = 0.7) +
  
  # Background all cities (small, gray)
  geom_point(size = 2, color = "lightgray", alpha = 0.3) +
  
  # Highlight extreme cities
  geom_point(
    data = extreme_data,
    aes(color = quadrant, size = population / 1e6, shape = coastal_status),
    alpha = 0.8,
    stroke = 1
  ) +
  
  # Label extreme cities only
  ggrepel::geom_text_repel(
    data = extreme_data,
    aes(label = city, color = quadrant),
    size = 3.5,
    max.overlaps = 24,
    segment.size = 0.4,
    segment.alpha = 0.5,
    box.padding = 0.4,
    fontface = "bold"
  ) +
  
  scale_color_manual(
    values = c(
      "High SIF\nHigh CO2" = "#e41a1c",
      "High SIF\nLow CO2" = "#4daf4a",
      "Low SIF\nHigh CO2" = "#ff7f00",
      "Low SIF\nLow CO2" = "#377eb8"
    ),
    name = "Quadrant"
  ) +
  
  scale_shape_manual(
    values = c("Coastal" = 17, "Inland" = 16),
    name = "Type"
  ) +
  
  scale_size_continuous(
    range = c(2, 8),
    name = "Population (M)"
  ) +
  
  labs(
    title = "All Cities: SIF Impact vs CO2 Concentration (All Seasons)",
    subtitle = "Extreme cities highlighted | ▲ = Coastal | ● = Inland | Gray = all other cities",
    x = "SIF",
    y = "CO2 Concentration"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )

print(p_combined)

```

```{r}
p_faceted <- ggplot(viz_data, aes(x = sif_iqr, y = xco2_iqr)) +
  
  # Quadrant dividing lines
  geom_hline(yintercept = median(viz_data$xco2_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.6, alpha = 0.7) +
  geom_vline(xintercept = median(viz_data$sif_iqr, na.rm = TRUE), 
             linetype = "dashed", color = "gray50", size = 0.6, alpha = 0.7) +
  
  # Background all cities
  geom_point(size = 1.5, color = "lightgray", alpha = 0.3) +
  
  # Highlight extreme cities
  geom_point(
    data = extreme_data,
    aes(color = quadrant, size = population / 1e6, shape = coastal_status),
    alpha = 0.8,
    stroke = 1
  ) +
  
  # Label extreme cities
  ggrepel::geom_text_repel(
    data = extreme_data,
    aes(label = city, color = quadrant),
    size = 3,
    max.overlaps = 20,
    segment.size = 0.3,
    segment.alpha = 0.5,
    box.padding = 0.3,
    fontface = "bold"
  ) +
  
  scale_color_manual(
    values = c(
      "High SIF\nHigh CO2" = "#e41a1c",
      "High SIF\nLow CO2" = "#4daf4a",
      "Low SIF\nHigh CO2" = "#ff7f00",
      "Low SIF\nLow CO2" = "#377eb8"
    ),
    name = "Quadrant"
  ) +
  
  scale_shape_manual(
    values = c("Coastal" = 17, "Inland" = 16),
    name = "Type"
  ) +
  
  scale_size_continuous(
    range = c(2, 6),
    name = "Population (M)"
  ) +
  
  facet_wrap(~season, ncol = 1) +
  
  labs(
    title = "Extreme Cities by Season: SIF Impact vs CO2 Concentration",
    subtitle = "2 columns x 5 rows | ▲ = Coastal | ● = Inland",
    x = "SIF Effect",
    y = "CO2 Concentration"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    strip.text = element_text(face = "bold", size = 10),
    panel.grid.minor = element_blank(),
    aspect.ratio = 1.2  # ← Make each panel taller
  )

options(repr.plot.height = 24)

print(p_faceted)
```

## Summary

-   53 locations, single temporal state (median for 2019-2023); co2 conc, co2 emissions, sif, population, gdp, coastal proximity

-   346 locations, by seasons / climate zones from 2019-2023 based; co2 conc, sif, population, gdp coastal proximity

## Notes (Apr 9)

Contributions:

-   Quadrant view in figma:

    -   Step 1 'Personas' of cities at the extremities of each quadrant

    -   Step 2: Adding bespoke variables to each city to consider adding to the analysis -\> localize analysis, spot patterns (regression/ pareto / tbd)

        -   may not be necessary for a relationship between variables to trigger Step 3. leave to cities to decide

    -   Step 3: \[Variable-to-action framework\] Introduce things that cities can do -\> Add possible policy actions to each variable (e.g. coastal proximity -\> harness wind? sif -\> nature-based solutions) -\> examples that I give

-   Concentration vs emissions -\> carbon impact score

    -   highlight a few cities

        -   MVP: 53 cities

        -   Stretch: from 346 cities if co2 emissions data can be expanded

Completeness of analysis

-   is there an ideal regression?

Looking at the cities in a quadrant, put into a framework for potential variables to investigate further

## Further work

-   Extend CO2 emission estimates from current 53 cities to expanded 346 locations

    -   check feasibility as data extraction and processing might be challenging due to uncertainty and data quality/ temporal resolution (ODIAC/ EDGAR)

-   Time of day as a variable (morning, afternoon)

-   Ways to measure CO2 differently as a dependent variable e.g. rate of change

-   Pareto-optimal
