---
title: "Carbon Cities"
format: 
  html:
    theme: cosmo
    toc: true
    code-fold: true
    self-contained: false 
    page-layout: full
editor: visual
warning: false
date: 04/05/2026
date-format: long
author: 
 - name: He Yutian
abstract: > 
  A study of 53 C40 cities, as part of "Speculative Carbon Futures Amidst Climate Data Inequities (working title)"
---

## Grid pixel heatmap demo (dash plotly)

-   Data exploration to identify trends; shows spatial-temporal spread of data, data gaps, and allows for more detailed sorting

-   Inverse relationship between CO2 and SIF - hypothesis about SIF's impact on CO2? Nature-based solutions

```{=html}
<iframe src="http://localhost:8052" height="700" width="100%" style="border:none; pointer-events: none;"></iframe>
```

## Visual Prototype Overview (Figma)

-   Building a tool to go from location-based selection of points of interest to comparing variables relevant to CO2

```{=html}
<iframe style="border: 1px solid rgba(0, 0, 0, 0.1);" width="100%" height="700" src="https://embed.figma.com/proto/nOUZp57ZbuHiR0M840E261/OCO3?node-id=553-4205&p=f&scaling=contain&content-scaling=fixed&page-id=0%3A1&starting-point-node-id=553%3A4205&show-proto-sidebar=1&embed-host=share" allowfullscreen></iframe>
```

## Visual prototype demo (js)

-   Demo of some of the Figma prototype functionalities using real data; WIP.

```{=html}
<iframe src="globe/index.html" width="100%" height="900px" style="border:none;"></iframe>
```

## Regression analysis

Carbon impact score (CO2 conc - CO2 emissions)

-   A 'carbon impact score' is introduced by taking the normalized interquartile range of CO2 concentration levels (OCO-3 netcdf data) subtracted by CO2 emissions levels (satellite-based estimates from Ahn et al, 2025).

-   Higher score = more "harmed" (high conc low emissions), lower score = more "harming" done (low conc high emissions)

-   This is then compared against solar induced fluorescence (SIF), population, gdp, number of samples, and coastal proximity via running a multivariate linear regression.

```{r}
# reference: https://towardsdatascience.com/7-steps-to-run-a-linear-regression-analysis-using-r-d5897a66b835/

# install.packages("tidyverse")
library(tidyverse)


df <- read.csv("c40cities-co2-em-conc-wkt-coast.csv") # outcome of https://github.com/heytian/d2d-oco3-tools/blob/main/20260405_04_wkt-coastal.py
df <- df |> drop_na(population)

# normalize independent variables using interquartile range 
iqr_norm <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  (x - q1) / (q3 - q1)
}


norm_df <- df |> 
  mutate(
    co2_conc_iqr = iqr_norm(xco2_ppm),
    co2_em_iqr = iqr_norm(annual_co2_em_mtco2),
    sif_iqr = iqr_norm(sif_757nm),
    pop_iqr = iqr_norm(population),
    xco2numsams_iqr = iqr_norm(xco2_numsams),
    sifnumsams_iqr = iqr_norm(sif_numsams),
    gdp_iqr = iqr_norm(gdp_bil_usd),
    coast_iqr = iqr_norm(coastal_km),
    impact_iqr = co2_conc_iqr - co2_em_iqr
  )
    

lm_iqr_impact<- lm(impact_iqr ~ sif_iqr + pop_iqr + xco2numsams_iqr + sifnumsams_iqr + gdp_iqr + coast_iqr, 
                  data = norm_df)
summary(lm_iqr_impact)
```

## Plots

This is the distribution of cities across the quadrant framework of high-low CO2 concentrations and emissions.

```{r}
# bubble size = population, color = quadrant, shape = coastal/inland
ggplot(norm_df, aes(x = co2_em_iqr, y = co2_conc_iqr)) +
  
  # quadrant dividing lines at IQR thresholds (0 = Q1, 1 = Q3)
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray60") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray60") +
  
  # cities as bubbles
  geom_point(aes(
    color = quadrant,
    size  = population / 1e6, # size by population millions
    shape = is_coastal
  ), alpha = 0.75) +
  
  # label interesting cities only
  ggrepel::geom_text_repel(
    data = norm_df |> filter(
      quadrant %in% c("LOW_emissions_HIGH_conc", "HIGH_emissions_LOW_conc")
    ),
    aes(label = city),
    size = 2.8, max.overlaps = 15
  ) +
  
  scale_color_manual(values = c(
    "HIGH_both"              = "#d73027",
    "LOW_both"               = "#4575b4",
    "HIGH_emissions_LOW_conc" = "#fc8d59",
    "LOW_emissions_HIGH_conc" = "#91bfdb")) +
  scale_shape_manual(values = c("True" = 17, "False" = 16),
                     labels = c("True" = "Coastal", "False" = "Inland")) +
  scale_size_continuous(range = c(2, 10), name = "Population (M)") +
  
  labs(
    title    = "City CO2 Concentration vs Emissions",
    subtitle = "Quadrants based on IQR thresholds | Size = population | Shape = coastal/inland",
    x        = "CO2 Emissions (IQR normalized)",
    y        = "CO2 Concentration (IQR normalized)",
    color    = "Quadrant",
    shape    = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

Impact score by quadrant, color differentiated by coastal proximity. Coastal cities cluster differently across the CO2 emissions - CO2 concentration quadrants

```{r}
ggplot(norm_df, aes(x = reorder(quadrant, impact_iqr, median), 
                     y = impact_iqr)) +
  geom_boxplot(outlier.shape = NA, fill = "gray90") +
  geom_jitter(aes(color = coastal_km, size = population / 1e6),
              width = 0.2, alpha = 0.8) +
  scale_color_viridis_c(
    name   = "Distance to\ncoast (km)",
    option = "viridis"
  ) +
  scale_size_continuous(range = c(2, 6), name = "Population (M)") +
  labs(
    title    = "Impact Score Distribution by Quadrant",
    subtitle = "Color = coastal distance | Size = population",
    x        = NULL,
    y        = "Impact Score (IQR normalized)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

Regression analysis shows that coastal distance and population are key drivers.

```{r}
library(tidyverse)
library(broom)

# tidy regression output
coef_df <- tidy(lm_iqr_impact, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(term = recode(term,
    "sif_iqr"         = "SIF",
    "pop_iqr"         = "Population",
    "xco2numsams_iqr" = "CO2 Sample Size",
    "sifnumsams_iqr"  = "SIF Sample Size",
    "gdp_iqr"         = "GDP",
    "coast_iqr"       = "Coastal Distance"
  ))

ggplot(coef_df, aes(x = estimate, y = reorder(term, estimate))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_point(aes(color = p.value < 0.20), size = 3) +
  scale_color_manual(
    values = c("FALSE" = "gray70", "TRUE" = "red"),
    labels = c("p > 0.20", "p < 0.20")
  ) +
  labs(
    title = "Predictors of CO2 Impact Score",
    subtitle = "Impact score = normalized CO2 concentration − normalized emissions",
    x = "Coefficient estimate (IQR normalized)",
    y = NULL,
    color = NULL
  ) +
  theme_minimal()
```

The coastal effect is most visible in cities with low CO2 emissions and high CO2 concentrations. Coastal proximity may partially explain why these cities end up receiving more CO2 despite emitting low amounts.

```{r}
ggplot(norm_df, aes(x = coastal_km, y = impact_iqr)) +
  geom_point(aes(color = quadrant, size = population / 1e6), alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "gray30", linewidth = 0.7) +
  scale_color_manual(values = c(
    "HIGH_both"               = "#d73027",
    "LOW_both"                = "#4575b4",
    "HIGH_emissions_LOW_conc" = "#fc8d59",
    "LOW_emissions_HIGH_conc" = "#91bfdb"
  )) +
  scale_size_continuous(range = c(2, 8), name = "Population (M)") +
  facet_wrap(~quadrant, scales = "free_y") +
  labs(
    title    = "Coastal Distance vs Impact Score by Quadrant",
    x        = "Distance to coast (km)",
    y        = "Impact Score (IQR normalized)",
    color    = "Quadrant"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Further work

-   Introduce temporal element to 53-cities analysis: add in time stamps, look at seasonality/ climate zone as another independent variable

-   Extend regression analysis from 53 to 320 cities

    -   CO2 conc & sif

    -   Feasibility of extending other variables to 320 cities:

        -   coastal proximity through join by nearest

        -   GDP through pro-rated population calc

        -   CO2 emissions from processing ODIAC/EDGAR (challenging due to uncertainty and data quality)

-   Ways to measure CO2 differently as a dependent variable e.g. rate of change
